```{r}
library(tidyverse)
```

```{r}
soccer <- read_csv("https://dwillis.github.io/sports-data-files/ncaa_womens_soccer_totals_2024.csv")
```

The Southwestern Athletic Conference is unrepresented in the national rankings and is often in the lower half of all Division I conferences by overall record. But on another list, the SWAC dominates.

Prairie View, Mississippi Valley State and Alabama A&M were the top three women's soccer teams in the country in 2024 by the percentage of their shot attempts that were on goal. Arkansas-Pine Bluff - another SWAC member - ranked sixth. 

Prairie View, which led the country in shot on goal percentage by over four percentage points, and Mississippi Valley State were the only teams to eclipse 60 percent.

Only three of the 78 teams in the Power Five conferences even surpassed 50 percent.

```{r}
powerfive <- c("Big Ten", "ACC", "Big 12", "Big East", "SEC")

soccer |> 
  filter(conference_name %in% powerfive) |> 
  mutate(SoGPct = (SoG/ShAtt)*100) |>
  arrange(desc(SoGPct))
```

