---
title: "Women's Soccer Data"
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1G1AjH2iS6kj3kKHCRi0XVy5Lcm-rahJkeTAToVh9J1g/edit?gid=0#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2025 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:


```{r}
matches_2025 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2025.csv") |> mutate(diff = team_score - opponent_score)
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2025)
glimpse(matches_2025)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1. How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal.

```{r}
matches_2025 |> 
  group_by(team) |> 
  summarise(
    total_shots = sum(sh_att),
    total_sog = sum(so_g),
    pct_sog = total_sog/total_shots
  ) |> 
arrange(desc(pct_sog))
```

2. What about the same shot stats from last season's first four games? You'll need to load the matches from 2024 - just change the season in the URL - and filter by date to find the first four games.

```{r}
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv")
```

```{r}
firstfourgames <- c("2024-08-15", "2024-08-18", "2024-08-22", "2024-08-25")

matches_2024 |> 
  filter(team == "Maryland Terrapins, Terps") |> 
  filter(date %in% firstfourgames) |> 
  group_by(team) |> 
  summarise(
    total_shots = sum(sh_att),
    total_sog = sum(so_g),
    pct_sog = total_sog/total_shots
  )

```



3. Maryland started 2022 with 4 draws. Have any other teams had as many this season?

```{r}
matches_2025 |> 
  group_by(team) |> 
  summarise(
    total_draws = sum(outcome == "Draw")
  ) |> 
arrange(desc(total_draws))
  
```


4. How many total corner kicks has Maryland had so far this season and how many have its opponents have? Calculate the difference between the two totals for every team using group_by, summarize and mutate, then filter to find Maryland's totals.

```{r}
matches_2025 |> 
  group_by(team) |>
  mutate(corners_diff = corners - defensive_corners) |> 
  summarise(total_corners_diff = sum(corners_diff)
              ) |> 
filter(team == "Maryland Terrapins, Terps")
```

```{r}
swac_teams <- c("Alabama A&M Bulldogs", "Alabama St. Lady Hornets", "Alcorn Lady Braves", "Ark.-Pine Bluff Lady Lions", "Grambling Lady Tigers", "Jackson St. Lady Tigers", "Mississippi Val. Devilettes", "Prairie View Lady Panthers", "Southern U. Lady Jaguars", "Texas Southern Lady Tigers")

swac_matches <- matches_2025 |> 
  filter(team %in% swac_teams)

matches_2025 |> 
  summarise(
    mean_diff = mean(diff),
    mean_opp_goals = mean(defensive_goals),
    mean_sog = mean(so_g),
    mean_opp_sog = mean(defensive_so_g)
    )

swac_matches |> 
  summarise(
    mean_diff = mean(diff),
    mean_opp_goals = mean(defensive_goals),
    mean_sog = mean(so_g),
    mean_opp_sog = mean(defensive_so_g)
    )

big10 <- c("Nebraska Huskers", "Iowa Hawkeyes", "Minnesota Golden Gophers", "Illinois Fighting Illini", "Northwestern Wildcats", "Wisconsin Badgers", "Indiana Hoosiers", "Purdue Boilermakers", "Ohio St. Buckeyes", "Michigan Wolverines", "Michigan St. Spartans", "Penn St. Nittany Lions", "Rutgers Scarlet Knights", "Maryland Terrapins, Terps", "Southern California Trojans", "UCLA Bruins", "Washington Huskies", "Oregon Ducks")

big10_matches <- matches_2025 |> 
  filter(team %in% big10)

big10_matches |> 
  summarise(
    mean_diff = mean(diff),
    mean_opp_goals = mean(defensive_goals),
    mean_sog = mean(so_g),
    mean_opp_sog = mean(defensive_so_g)
    )


big10_matches |> 
  filter(team == "Maryland Terrapins, Terps") |> 
  summarise(
    mean_diff = mean(diff),
    mean_opp_goals = mean(defensive_goals),
    mean_sog = mean(so_g),
    mean_opp_sog = mean(defensive_so_g)
    )
```


5. What's the most interesting thing you've seen from these questions?

The most newsworthy statistic I've noticed from these questions is the considerable disparity in shot attempts, shots on goal, and shots on goal percentage for Maryland women's soccer over the first four games of the 2024 season compared to the 2025 season. They've attempted more than double the shots and tallied over twice as many shots on goal this season than they did in 2024. Their shots on goal percentage also increased by two points. 

This could be strong and engaging statistical evidence in a story about the overall difference between the team last year compared to this year. Or, it could illustrate how much more effective the offense has been in head coach Michael Marchiano's first year leading the program.