
```{r}
library(tidyverse)
library(janitor)
library(ggplot2)
```

I have this data for every Super Bowl team since 2000, which is the timeframe I'll be focusing on. I'll have to rename a few of the columns that repeat, since this dataframe includes rushing and receiving stats, so "TD" and "Succ%" are repeated as column names. I plan on looking at the rushing yards per team (and how that has changed over time), and how many Super Bowl-winning teams had a 1,000-yard rusher. 

ProFootballReference also has a section that lists a team's league ranking in several rushing statistics (Attempts, Yards, Yards Per Carry, Touchdowns), so I could do a bar chart or line chart showing the league placements of every Super Bowl team since 2000.

```{r}
eagleslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2024eagles.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2024 Eagles")
eaglesplayoffslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2024eaglesplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2024 Eagles")
eaglesleaguerank2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2024eaglesleaguerank.csv") |> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2024 Eagles")
```

```{r}
chiefslogs3 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2023chiefs.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch)  |> mutate(Team = "2023 Chiefs")
chiefsplayoffslogs3 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2023chiefsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch)  |> mutate(Team = "2023 Chiefs")
chiefsleaguerank3 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2023chiefsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2023 Chiefs")
```

```{r}
chiefslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2022chiefs.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |>  mutate(Team = "2022 Chiefs")
chiefsplayoffslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2022chiefsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2022 Chiefs")
chiefsleaguerank2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2022chiefsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2022 Chiefs")
```

```{r}
ramslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2021rams.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2021 Rams")
ramsplayoffslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2021ramsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2021 Rams")
ramsleaguerank <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2021ramsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2021 Rams")
```

```{r}
bucslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2020bucs.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2020 Bucs")
bucsplayoffslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2020bucsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2020 Bucs")
bucsleaguerank2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2020bucsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2020 Bucs")
```

```{r}
chiefslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2019chiefs.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2019 Chiefs")
chiefsplayoffslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2019chiefsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2019 Chiefs")
chiefsleaguerank1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2019chiefsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2019 Chiefs")
```

```{r}
patriotslogs6 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2018patriots.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2018 Patriots")
patriotsplayoffslogs6 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2018patriotsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2018 Patriots")
patriotsleaguerank6 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2018patriotsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2018 Patriots")
```

```{r}
eagleslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2017eagles.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2017 Eagles")
eaglesplayoffslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2017eaglesplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2017 Eagles")
eaglesleaguerank1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2017eaglesleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2017 Eagles")
```

```{r}
patriotslogs5 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2016patriots.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2016 Patriots")
patriotsplayoffslogs5 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2016patriotsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2016 Patriots")
patriotsleaguerank5 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2016patriotsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2016 Patriots")
```

```{r}
broncoslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2015broncos.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2015 Broncos")
broncosplayoffslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2015broncosplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2015 Broncos")
broncosleaguerank <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2015broncosleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2015 Broncos")
```

```{r}
patriotslogs4 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2014patriots.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2014 Patriots")
patriotsplayoffslogs4 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2014patriotsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2014 Patriots")
patriotsleaguerank4 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2014patriotsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2014 Patriots")
```

```{r}
seahawkslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2013seahawks.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2013 Seahawks")
seahawksplayoffslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2013seahawksplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2013 Seahawks")
seahawksleaguerank <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2013seahawksleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2013 Seahawks")
```

```{r}
ravenslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2012ravens.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2012 Ravens")
ravensplayoffslogs2<- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2012ravensplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2012 Ravens")
ravensleaguerank2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2012ravensleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2012 Ravens")
```

```{r}
giantslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2011giants.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2011 Giants")
giantsplayoffslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2011giantsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2011 Giants")
giantsleaguerank2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2011giantsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2011 Giants")
```

```{r}
packerslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2010packers.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2010 Packers")
packersplayoffslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2010packersplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2010 Packers")
packersleaguerank <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2010packersleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2010 Packers")
```

```{r}
saintslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2009saints.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2009 Saints")
saintsplayoffslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2009saintsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2009 Saints")
saintsleaguerank <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2009saintsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2009 Saints")
```

```{r}
steelerslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2008steelers.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2008 Steelers")
steelersplayoffslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2008steelersplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2008 Steelers")
steelersleaguerank2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2008steelersleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2008 Steelers")
```

```{r}
giantslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2007giants.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2007 Giants")
giantsplayoffslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2007giantsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2007 Giants")
giantsleaguerank1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2007giantsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2007 Giants")
```

```{r}
coltslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2006colts.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2006 Colts")
coltsplayoffslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2006coltsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2006 Colts")
coltsleaguerank <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2006coltsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2006 Colts")
```

```{r}
steelerslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2005steelers.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2005 Steelers")
steelersplayoffslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2005steelersplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2005 Steelers")
steelersleaguerank1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2005steelersleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2005 Steelers")
```

```{r}
patriotslogs3 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2004patriots.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2004 Patriots")
patriotsplayoffslogs3 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2004patriotsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2004 Patriots")
patriotsleaguerank3 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2004patriotsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2004 Patriots")
```

```{r}
patriotslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2003patriots.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2003 Patriots")
patriotsplayoffslogs2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2003patriotsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2003 Patriots")
patriotsleaguerank2 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2003patriotsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2003 Patriots")
```

```{r}
bucslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2002bucs.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2002 Bucs")
bucsplayoffslogs <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2002bucsplayoffs.csv") |> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2002 Bucs")
bucsleaguerank1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2002bucsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2002 Bucs")
```

```{r}
patriotslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2001patriots.csv") |> clean_names() |> rename ("RushTD" = "td_9") |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2001 Patriots")
patriotsplayoffslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2001patriotsplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2001 Patriots")
patriotsleaguerank1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2001patriotsleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2001 Patriots")
```

```{r}
ravenslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2000ravens.csv") |> clean_names()|> rename ("RushTD" = "td_9") |> rename("yds" = `yds_8`) |> filter(pos == "RB") |> select(player, att, yds, y_a, RushTD, y_scm, rrtd, touch) |> mutate(Team = "2000 Ravens")
ravensplayoffslogs1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2000ravensplayoffs.csv")|> rename("Yards" = `Yds...8`) |> rename ("RushTD" = `TD...9`) |> rename("YPC" = "Y/A") |> filter(Pos == "RB") |> select(Player, Att, Yards, YPC, RushTD, YScm, RRTD, Touch) |> mutate(Team = "2000 Ravens")
ravensleaguerank1 <- read_csv("~/GitHub/jour479x_fall_2025/presentations/Presentation2/2000ravensleaguerank.csv")|> rename("RushAttempts" = "Att...16") |> rename("RushYards" = `Yds...17`) |> rename("RushTDs" = `TD...18`) |> rename("YPC" = "Y/A") |>  filter(Player == "Lg Rank Offense") |> select(RushAttempts, RushYards, RushTDs, YPC) |> mutate(Team = "2000 Ravens")
```

```{r}
logs <- rbind(eagleslogs2, chiefslogs3, chiefslogs2, ramslogs, bucslogs2, chiefslogs1, patriotslogs6, eagleslogs1, patriotslogs5, broncoslogs, patriotslogs4, seahawkslogs, ravenslogs2, giantslogs2, packerslogs, saintslogs, steelerslogs2, giantslogs1, coltslogs, steelerslogs1, patriotslogs3, patriotslogs2, bucslogs1, patriotslogs1, ravenslogs1)
```

```{r}
playoffslogs <- rbind(eaglesplayoffslogs2, chiefsplayoffslogs3, chiefsplayoffslogs2, ramsplayoffslogs, bucsplayoffslogs2, chiefsplayoffslogs1, patriotsplayoffslogs6, eaglesplayoffslogs1, patriotsplayoffslogs5, broncosplayoffslogs, patriotsplayoffslogs4, seahawksplayoffslogs, ravensplayoffslogs2, giantsplayoffslogs2, packersplayoffslogs, saintsplayoffslogs, steelersplayoffslogs2, giantsplayoffslogs1, coltsplayoffslogs, steelersplayoffslogs1, patriotsplayoffslogs3, patriotsplayoffslogs2, bucsplayoffslogs, patriotsplayoffslogs1, ravensplayoffslogs1)
```

```{r}
leagueranklogs <- rbind(eaglesleaguerank2, chiefsleaguerank3, chiefsleaguerank2, ramsleaguerank, bucsleaguerank2, chiefsleaguerank1, patriotsleaguerank6, eaglesleaguerank1, patriotsleaguerank5, broncosleaguerank, patriotsleaguerank4, seahawksleaguerank, ravensleaguerank2, giantsleaguerank2, packersleaguerank, saintsleaguerank, steelersleaguerank2, giantsleaguerank1, coltsleaguerank, steelersleaguerank1, patriotsleaguerank3, patriotsleaguerank2, bucsleaguerank1, patriotsleaguerank1, ravensleaguerank1)
```

```{r}
leagueranklogs <- leagueranklogs |> mutate(Mean = mean(RushYards))
```


My initial question has not changed, though I've narrowed my scope to only the NFL, rather than also including college football. Just from a cursory glance at the data, it seems that elite running backs are not necessary to winning a Super Bowl, but I'd like to put numbers to this claim. 15 of the past 25 Super Bowl champions did not have a 1,000 yard rusher, which I will express in my presentation.

For the season-long rushing numbers and playoff rushing statistics, I plan on visualizing those using a bar chart or line chart, then contextualizing that data with the league-wide ranking. So for the 2024 Eagles, I would show their running backs' statistics from the regular season and postseason, then show that they were near the top of the league in many rushing metrics.

```{r}
ggplot() +
  geom_bar(
    data = leagueranklogs,
    aes(
      x = Team,
      weight = RushYards
    )
  ) + geom_hline(yintercept = 14.92, linetype = "dashed") +
  coord_flip() +
  labs(
    title = "Super Bowl Champions' Rushing Production Has Varied",
    subtitle = "Since 2000, Super Bowl champions have ranked roughly in the middle of the pack in rushing \nyards on average",
    x = "Team",
    y = "Total Rushing Yards League Ranking"
  )
```

The only question I have as of right now is about creating a beeswarm plot showing the number of Super Bowl champions in each range of league ranking (how many ranked from first to fifth in the league in rushing yards, how many were ranked between sixth and 10th, etc.). The x-axis of the beeswarm plot would default to displaying every individual ranking, so I need to create a column that designates which team falls in which range, I'm just unsure of how to do that.

For the bar chart I created, I'd like to label the dotted line (representing the average league ranking of every Super Bowl team since 2000) to clarify what it is and what the exact value is. Would a legend or an annotation in the chart work better? If a legend would work, how do I create a legend that doesn't overlap with the chart?

I also have a question that's less about syntax or RStudio itself, but about your advice on what would be best to include in the project. Do you think I should create a data visualization showing the raw rushing numbers (yards, TDs, yards per carry, etc.) of the Super Bowl-winning teams, or does their league ranking illustrate how productive they were on the ground? 

